---
title: "CD14+ Monocytes"
author: "Chih-Hsuan Wu"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
source("./code/hippo_all_proportion.R")
source("./code/DE_methods.R")
source("./code/functions_in_rmd.R")
load("./data/data2sce.Rdata")
# load DE results generated from "analysis on Kang.Rmd"
load("./data/10X_Kang_DEresult.RData")
library(ggplot2)
library(dplyr)
library(ggpubr)
library(SingleCellExperiment)
library(pheatmap)
library(gridExtra)
library(reshape)
library(clusterProfiler)
library(enrichplot)
library("org.Hs.eg.db")
library(pathview)
assign_celltype("CD14+ Monocytes")
subgroupsce$ind = paste0("D",subgroupsce$ind)
Kang_Wilcox_Seurat_df[[key]]$avg_log2FC = -Kang_Wilcox_Seurat_df[[key]]$avg_log2FC
Kang_Wilcox_Seurat_df[[key]]$genes = gsub("-", "_", rownames(Kang_Wilcox_Seurat_df[[key]]))
for(i in 1:length(Kang_Wilcox_Seurat_df[[key]]$genes)){
  if(Kang_Wilcox_Seurat_df[[key]]$genes[i] %in%rownames(subgroupsce)){}else{
    Kang_Wilcox_Seurat_df[[key]]$genes[i] = gsub("_ENSG", "-ENSG", Kang_Wilcox_Seurat_df[[key]]$genes[i])
  }
}
```

```{r}
group_palette = c("#00BCD4", "#E91E63")
celltype_palette = setNames(c( "#6A00FF", "#F0A30A"), c("CD14+ Monocytes","B cells"))
# , "#A4C400", "#6D8764", "#F472D0",
  # "#76608A", "#A20025", "#00ABA9", "#A0522D", "#008A00",
  # "#647687", "#1BA1E2", "#FA6800", "#825A2C", "#0050EF",
  # "#E3C800", "#E51400", "#D80073", "#60A917", "#AA00FF")
donors_palette = c("#f8766d", "#ce9700", "#7dae00", "#01be67", "#01bfc4", "#01a9ff", "#c77cff", "#ff61cc")
annotation_colors = list(Group = setNames(group_palette, c("Ctrl", "Stim")), Donor = setNames(donors_palette, sort(unique(subgroupsce$ind))))
```

# Data summary

```{r}
df = data.frame(cellgroup = subgroupsce$group_id, donor = subgroupsce$ind, count = 1)
df = aggregate(df$count, list(df$donor, df$cellgroup), FUN = sum)
colnames(df) = c("Donor", "Group", "Count")
df = df[nrow(df):1,]
df$cumcount= c(t(aggregate(df$Count, list(rev(df$Group)), FUN = cumsum)[,-1]))
p_donor = ggplot(df, aes(x = Group, y = Count, fill = Donor)) + 
  geom_bar(stat = "identity") + 
  xlab("Group") + 
  ylab("Number of cells") + 
  scale_fill_manual(values = donors_palette) +
  # labs(fill = "donors", labels=c("D7", "D6", "D5", "D4", "D3")) + 
  geom_text(aes(label=Count, y = cumcount), vjust = -0.2, size = 2) + 
  theme_bw() + 
  theme(panel.grid = element_blank())

```

```{r, eval = FALSE}
composition = data.frame(
  zeromean = sum(glm_df[[key]]$status=="zero mean"),
  lowlyexpressed = sum(glm_df[[key]]$status=="lowly expressed"), 
  valid = sum(glm_df[[key]]$status=="done"))
composition = melt(composition)

p_composition = ggplot(composition, aes(x = "", y = value, fill = variable)) + 
  geom_bar(stat = "identity") + 
  xlab("") + 
  ylab("number of genes") + 
  scale_fill_discrete(name = "", labels=c("zero mean", "freq < 5%", "freq > 5%")) +
  geom_text(aes(label=value), position = position_stack(vjust = 0.5), size = 3) + 
  theme_bw() + theme(panel.grid = element_blank()) + 
  ggtitle(paste("gene composition of ", key))
```

```{r}
# gene count per cell by group
count_cell_df = data.frame(totalcount = apply(subgroupsce@assays@data$counts, 2, sum), Group = subgroupsce$group_id, Donor = subgroupsce$ind)

p_countbygroup = ggplot(count_cell_df, aes(x = totalcount, color = Group)) +
  geom_histogram(aes(y=after_stat(density)), fill= "white", alpha=0.5, position="identity") +
  geom_density(alpha = 0.2) +
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  scale_colour_manual(labels = c("Ctrl", "Stim"), values = group_palette) +
  xlim(c(0,5000))
```
```{r}
# gene count per cell by donor
p_countbydonor = ggplot(count_cell_df, aes(x = totalcount, color = Donor)) +
  geom_histogram(aes(y=after_stat(density)),fill="white", alpha=0.5, position="identity") + 
  geom_density(alpha = 0.2) + 
  xlab("Library size") +
  ylab("Density") +
  theme_bw() + theme(panel.grid = element_blank()) + 
  scale_color_manual(values = donors_palette) + 
  xlim(c(0,5000))
```
```{r}
# zero proportion per gene
zeroprop_df = rbind(data.frame(zero_prop = rowMeans(subgroupsce@assays@data$counts[, subgroupsce$group_id%in%group1]== 0), genemean = rowMeans(subgroupsce@assays@data$counts[, subgroupsce$group_id%in%group1]), Group = factor(group1)),
                data.frame(zero_prop = rowMeans(subgroupsce@assays@data$counts[, subgroupsce$group_id%in%group2]== 0), genemean = rowMeans(subgroupsce@assays@data$counts[, subgroupsce$group_id%in%group2]), Group = factor(group2)),
                data.frame(zero_prop = rowMeans(subgroupsce@assays@data$counts== 0), genemean = rowMeans(subgroupsce@assays@data$counts), Group = factor("All B cells"))
                )
zeroprop_df = zeroprop_df[which(zeroprop_df$genemean>0),]
levels(zeroprop_df$Group) = c("Ctrl", "Stim", "All B cells")
p_zeroprop = ggplot(zeroprop_df, aes(x = genemean, y = zero_prop)) + 
  geom_point(aes(shape = Group, color = Group), alpha=0.5, size = 1.2) + 
  xlab("Gene mean") + 
  ylab("Zero proportion") +
  scale_color_manual(values = c(group_palette,"darkmagenta")) + 
  scale_shape_manual(values = c(16,16,3)) +
  xlim(c(0,10)) + 
  stat_function(fun = function(x) exp(-x), colour = "black") +
  theme_bw() + theme(panel.grid = element_blank())

```

```{r, fig.width=8}
ggarrange(p_donor, p_zeroprop, p_countbygroup, p_countbydonor, nrow = 2, ncol = 2)
```

# Number of DEGs from each method
```{r}
dflist = list(Kang_pois_glmm_df, 
              Kang_binom_glmm_df, 
              Kang_DESeq2.muscat,
              Kang_edgeR.cpm,
              Kang_MAST_df, 
              Kang_Wilcox_Seurat_df,
              # Kang_MMdream,
              Kang_MMvst,
              Kang_MMpoisson)
titlelist = c("Poisson-glmm_UMI", 
              "Binomial-glmm_UMI",
              "pb-DESeq2_UMI", 
              "pb-edgeR_cpm",
              "MAST_CPM",
              "Wilcox_int",
              # "MMdream",
              "MMvst_VST",
              "MMpoisson_UMI")


number_input_hits = data.frame(DEGs = integer(0), Input = integer(0))
pois_glmm_inputgene = Kang_pois_glmm_df[[key]]$genes[!is.na(Kang_pois_glmm_df[[key]]$hits)]
for (df in dflist){
  number_input_hits = rbind(number_input_hits, 
                            c(sum(!is.na(df[[key]]$hits)), 
                              sum(df[[key]]$hits, na.rm = T))) 
}
colnames(number_input_hits) = c("Inputs", "DEGs")
number_input_hits$method = titlelist

number_input_hits = melt(number_input_hits, id = "method")
colnames(number_input_hits)[2:3] = c("Gene", "Number_of_genes")
```


```{r}
ggplot(data=number_input_hits, aes(x=factor(method, 
                                            levels = titlelist), y=Number_of_genes, fill = Gene)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label=Number_of_genes), vjust=-0.2, position=position_dodge(0.9), size = 2) +
  theme_bw() + theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  scale_fill_manual(values = c("#BCAAA4", "#5D4037")) +
  xlab("") +
  ylab("Number of genes") +
  ggtitle(paste("DEGs summary of ", key))
```

To compare the results fairly, the genes displayed in all of the plots below are restricted to the inputs of pois_glmm. That is, the genes not containing in the inputs of pois_glmm are excluded.

```{r}
for (i in 1:length(dflist)){
  dflist[[i]][[key]] = dflist[[i]][[key]][dflist[[i]][[key]]$genes%in%pois_glmm_inputgene,]
}
for (df in dflist){
  df[[key]] = df[[key]][df[[key]]$genes%in%pois_glmm_inputgene,]
}
```
```{r}
number_input_hits = data.frame(DEGs = integer(0), input = integer(0))
pois_glmm_inputgene = Kang_pois_glmm_df[[key]]$genes[!is.na(Kang_pois_glmm_df[[key]]$hits)]
for (df in dflist){
  number_input_hits = rbind(number_input_hits, 
                            c(sum(!is.na(df[[key]]$hits)), 
                              sum(df[[key]]$hits, na.rm = T)))
}
colnames(number_input_hits) = c("Inputs", "DEGs")
number_input_hits$method = titlelist

number_input_hits = melt(number_input_hits, id = "method")
colnames(number_input_hits)[2:3] = c("Gene", "Number_of_genes")
```

```{r}
ggplot(data=number_input_hits, aes(x=factor(method, 
                                            levels = titlelist), y=Number_of_genes, fill = Gene)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label=Number_of_genes), vjust=-0.2, position=position_dodge(0.9), size = 2) +
  theme_bw() + theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  scale_fill_manual(values = c("#BCAAA4", "#5D4037")) +
  xlab("") +
  ylab("Number of genes") +
  ggtitle(paste("DEGs summary of ", key))
```

# Volcano plot
```{r, fig.width = 10, fig.height=6}
pvalcutoff = 0.05
v = 2
log2FCcutoff = log2(v)

p = list()
for (i in 1:length(dflist)){
  if(is.null(dflist[[i]][[key]]$genes)){
    labgene = rownames(dflist[[i]][[key]])
  }else{
    labgene = dflist[[i]][[key]]$genes
  }
  if(i %in% 6){
    xvar = 'avg_log2FC'
    yvar = 'p_val_adj'
  }else{
    xvar = 'log2FC'
    yvar = 'BH'
  }
  if (nrow(dflist[[i]][[key]])==0){next}
  p[[i]] = ggplot(na.omit(data.frame(dflist[[i]][[key]])), aes_string(x = xvar, y = sprintf("-log10(%s)", yvar), colour = "hits")) +
    geom_point(alpha = 0.5, size = 0.5) +
    labs(color = "hits") +
    geom_hline(yintercept=-log10(pvalcutoff),linetype=2) +
    geom_vline(xintercept=log2FCcutoff,linetype=2) + 
    geom_vline(xintercept=-log2FCcutoff,linetype=2) +
    ggtitle(titlelist[i]) + 
    theme_bw() + 
    theme(panel.grid = element_blank()) + 
    scale_color_manual(values = c("gray", "blue")) +
    xlim(c(-3,3)) + 
    ylim(c(0,50)) + 
    labs(color = "DEGs") +
    xlab("Log2 Fold Change") + 
    ylab("-Log10 (adj.pval)")
}

figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)

# Annotate the figure by adding a common labels
annotate_figure(figure,
                top = paste("Volcano plot",key),
                bottom = text_grob(paste( "log2FCcutoff", paste0("log2(",v,")"),"; adj.pval cutoff", pvalcutoff)
                                   , x = 0.99, just = "right")
)
```

# Histogram of p-value/adj.p-value
```{r, fig.width = 10, fig.height=6}
p = list()
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  df = data.frame(genes = dflist[[i]][[key]]$genes)
  if(i %in% 6){
    df = cbind(df, dflist[[i]][[key]][,c("p_val", "p_val_adj")])
  }else{
    df = cbind(df, dflist[[i]][[key]][,c("pval", "BH")])
  }
  
  p[[i]] = ggplot(melt(df,id = "genes"), aes(x=value, color=variable)) +
geom_histogram(fill="white", alpha=0.5, position="identity") + theme_bw() + theme(panel.grid = element_blank()) + xlab(NULL) + ylab("Count") + ylim(c(0,2500)) + ggtitle(titlelist[i]) + scale_colour_discrete(guide = guide_legend(title = NULL),labels = c("p-value", "adj.p-value"))
}
figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)


annotate_figure(figure, top = as.character(key))
```

# P-Value comparison across different methods
```{r,fig.width = 10, fig.height=6}
value_comparison("pval")
```

# Log2 fold change comparison across different methods
```{r,fig.width = 10, fig.height=6}
value_comparison("log2FC")
```

# Violin plot of log2mean of DEGs
```{r, fig.width=7}
p = list()
df = data.frame()
clusterlist = list(group1, group2)
clustername = c(paste(group1, collapse = "_"), paste(group2, collapse = "_"))
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  if (sum(dflist[[i]][[key]]$hits, na.rm = T)==0){next}
  if(is.null(dflist[[i]][[key]]$genes)){
    hits_gene = rownames(dflist[[i]][[key]])[which(dflist[[i]][[key]]$hits)]
  }else{
    hits_gene = dflist[[i]][[key]]$genes[which(dflist[[i]][[key]]$hits)]
  }
  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:2){
    cellind = subgroupsce$group_id %in% clusterlist[[j]]
    df = rbind(df, data.frame(Log2mean = log2(fun(subgroupsce@assays@data$counts[hits_gene,cellind])), Group = clustername[j], method = titlelist[i]))
  }
}
levels(df$Group) = clustername
```


```{r, fig.width=7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=Log2mean, fill = Group)) +
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylim(c(-3,4)) + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  scale_fill_manual(labels = c("Ctrl", "Stim"),values = group_palette)

```

# Violin plot of gene expression frequency of DEGs
````{r, fig.width = 7}
p = list()
df = data.frame()
clusterlist = list(key, levels(sce$cluster_id))
clustername = c(key, "All clusters")
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  if (sum(dflist[[i]][[key]]$hits, na.rm = T)==0){next}
  if(is.null(dflist[[i]][[key]]$genes)){
    hits_gene = rownames(dflist[[i]][[key]])[which(dflist[[i]][[key]]$hits)]
  }else{
    hits_gene = dflist[[i]][[key]]$genes[which(dflist[[i]][[key]]$hits)]
  }
  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:2){
    cellind = sce$cluster_id %in% clusterlist[[j]]
    df = rbind(df, data.frame(freq = fun(sce@assays@data$counts[hits_gene,cellind]>0), cluster = clustername[j], method = titlelist[i]))
  }
}
levels(df$cluster) = clustername
```

````{r, fig.width = 7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=freq, fill = factor(cluster, levels =  levels(cluster)))) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylab("Gene expression frequency") + 
  ylim(c(0,1)) + 
  xlab("") + 
  scale_fill_manual(values = setNames(c(celltype_palette[key], "gray"), c(key, "All clusters"))) +
  labs(fill = "Cluster") +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) 
```

# Heatmap of top DEGs

## Poisson-glmm DEGs
```{r}
sort_log2FC = sort(abs(Kang_pois_glmm_df[[key]]$log2FC[Kang_pois_glmm_df[[key]]$hits]), index.return = T, decreasing = T)
hitgenes = Kang_pois_glmm_df[[key]]$genes[which(Kang_pois_glmm_df[[key]]$hits)][sort_log2FC$ix]
mat = subgroupsce@assays@data$counts[hitgenes,]
#row annotation
annotation_log2FC = data.frame(abs_log2FC = abs(Kang_pois_glmm_df[[key]]$log2FC[match(hitgenes,Kang_pois_glmm_df[[key]]$genes)]))
rownames(annotation_log2FC) = rownames(mat)
#col annotation
annotation_df = data.frame(Donor = subgroupsce$ind, Group = subgroupsce$group_id)
levels(annotation_df$Group) = c("Ctrl", "Stim")
rownames(annotation_df) = colnames(mat)
annotation_df = annotation_df[with(annotation_df, order(Group, Donor)), ]
# cell level
pheatmap(mat[,rownames(annotation_df)],
         main = paste0("Poisson-glmm DEGs\nin UMI counts (", nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```

## Additional DEGs from other methods

### pb-DESeq2
```{r}
sort_log2FC = sort(abs(Kang_DESeq2.muscat[[key]]$log2FC[Kang_DESeq2.muscat[[key]]$hits]), index.return = T, decreasing = T)
hitgenes_deseq2 = Kang_DESeq2.muscat[[key]]$genes[which(Kang_DESeq2.muscat[[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_deseq2[!hitgenes_deseq2%in%hitgenes][1:600])
mat = subgroupsce@assays@data$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("pb-DESeq2 DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```

### MMpoisson
```{r}
sort_log2FC = sort(abs(dflist[[8]][[key]]$log2FC[dflist[[8]][[key]]$hits]), index.return = T, decreasing = T)
hitgenes_MMpoisson = dflist[[8]][[key]]$genes[which(dflist[[8]][[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_MMpoisson[!hitgenes_MMpoisson%in%hitgenes][1:600])
mat = subgroupsce@assays@data$counts[hits_notin_glmm,]


# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("MMpoisson DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```

# MA plot 
```{r fig.width = 8, fig.height=5}
fig_FC_mean = FC_mean_plot(subgroupsce@metadata$log2mean, c(-6,6), "Log2 mean")
annotate_figure(fig_FC_mean, top = paste("Log2FC vs. Log2 mean ",key))

```

# Enrichment analysis

## Poisson-glmm DEGs

### All DEGs
```{r}
original_gene_list = Kang_pois_glmm_df[[key]]$log2FC
names(original_gene_list) = Kang_pois_glmm_df[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
genes = Kang_pois_glmm_df[[key]]$genes[which(Kang_pois_glmm_df[[key]]$hits)]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```


### Up-regulated DEGs
```{r}
original_gene_list = Kang_pois_glmm_df[[key]]$log2FC
names(original_gene_list) = Kang_pois_glmm_df[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
positive_genes = which(Kang_pois_glmm_df[[key]]$hits & Kang_pois_glmm_df[[key]]$log2FC > 0)
genes = Kang_pois_glmm_df[[key]]$genes[positive_genes]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r, fig.width = 8, fig.height=6}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```

## pb-DESeq2 DEGs

### All DEGs
```{r}
original_gene_list = Kang_DESeq2.muscat[[key]]$log2FC
names(original_gene_list) = Kang_DESeq2.muscat[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
genes = Kang_DESeq2.muscat[[key]]$genes[which(Kang_DESeq2.muscat[[key]]$hits)]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```

### Up-regulated DEGs
```{r}
original_gene_list = Kang_DESeq2.muscat[[key]]$log2FC
names(original_gene_list) = Kang_DESeq2.muscat[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
positive_genes = which(Kang_DESeq2.muscat[[key]]$hits & Kang_DESeq2.muscat[[key]]$log2FC > 0)
genes = Kang_DESeq2.muscat[[key]]$genes[positive_genes]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r, fig.width = 8, fig.height=6}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```

## MMpoisson DEGs

### All DEGs
```{r}
original_gene_list = Kang_MMpoisson[[key]]$log2FC
names(original_gene_list) = Kang_MMpoisson[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
genes = Kang_MMpoisson[[key]]$genes[which(Kang_MMpoisson[[key]]$hits)]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```

### Up-regulated DEGs
```{r}
original_gene_list = Kang_MMpoisson[[key]]$log2FC
names(original_gene_list) = Kang_MMpoisson[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
positive_genes = which(Kang_MMpoisson[[key]]$hits & Kang_MMpoisson[[key]]$log2FC > 0)
genes = Kang_MMpoisson[[key]]$genes[positive_genes]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r, fig.width = 8, fig.height=6}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```

