---
title: "Dorsal Horn and Ventral Lateral White DE analysis"
author: "Chih-Hsuan Wu"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
source("./code/hippo_all_proportion.R")
source("./code/DE_methods.R")
source("./code/functions_in_rmd.R")
# load DE result generated from "analysis_humanspine.Rmd"
load("./data/human_spine_DEresult.RData")
load("./data/human_spine.RData")
load("./data/fallopian_tubes.RData")
library(ggplot2)
library(dplyr)
library(ggpubr)
library(SingleCellExperiment)
library(pheatmap)
library(gridExtra)
library(reshape)
library(clusterProfiler)
library(enrichplot)
library("org.Hs.eg.db")
library(pathview)

group1 = "Dors_Horn"
group2 = "Vent_Lat_White"
pair = c(group1, group2)
subgroupsce = sce_human_spine_filtered[, sce_human_spine_filtered$region%in%pair]
subgroup_Seurat = Seurat_human_spine_integrate[, Seurat_human_spine_integrate$region%in%pair]
cellgroup = ifelse(subgroupsce$region%in% group1, 1, 2)
  genemean1 = rowMeans(counts(subgroupsce)[,cellgroup==1])
  genemean2 = rowMeans(counts(subgroupsce)[,cellgroup==2])
  subgroupsce@metadata$log2mean = log2(genemean1*genemean2)/2
  subgroupsce@metadata$log2meandiff = log2(abs(genemean1-genemean2))

Wilcox_Seurat_df$avg_log2FC = - Wilcox_Seurat_df$avg_log2FC
```

```{r}
# old criteria
pois_glmm_df$hits = identifyDEGs(pois_glmm_df$BH, pois_glmm_df$log2FC, newcriteria = F)
binom_glmm_df$hits = identifyDEGs(binom_glmm_df$BH, binom_glmm_df$log2FC, newcriteria = F)
```



```{r}
region_palette = c(
  "#A0522D", "#6A00FF", "#A4C400", "#F472D0",
  "#0050EF", "#00ABA9", "#F0A30A", "#008A00",
  "#647687", "#1BA1E2", "#FA6800", "#E3C800")
donors_palette <- c("#FDA4AF", "#FCD34D", "#BEF264", "#67E8F9", "#A5B4FC", 
                    "#FF6F61", "#6A4C93", "#001688", "#009A33")
names(region_palette) = levels(as.factor(sce_human_spine_filtered$region))
names(donors_palette) = levels(as.factor(sce_human_spine_filtered$patient_ID))
annotation_colors = list(Region = region_palette[c(group1, group2)], Donor = donors_palette)
remove_y = theme(
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.title.y = element_blank()
)
```

# Data summary
```{r}
df = data.frame(Region = subgroupsce$region, Donor = subgroupsce$patient_ID, Count = 1)
df = aggregate(df$Count, list(df$Donor, df$Region), FUN = sum)
colnames(df) = c("Patient", "Region", "Count")
df = df[nrow(df):1,]
df$cumcount= c(t(aggregate(df$Count, list(rev(df$Region)), FUN = cumsum)[,-1]))
p_donor = ggplot(df, aes(x = Region, y = Count, fill = Donor)) + 
  geom_bar(stat = "identity") + 
  xlab("Region") + 
  ylab("Number of cells") + 
  scale_fill_manual(values = donors_palette) + 
  geom_text(aes(label=Count, y = cumcount), vjust = -0.2, size = 2) + 
  theme_bw() + 
  theme(panel.grid = element_blank())

```

```{r}
# gene count per cell by region
count_cell_df = data.frame(totalcount = apply(assays(subgroupsce)$counts, 2, sum), Region = as.factor(subgroupsce$region), Donor = as.factor(subgroupsce$patient_ID), Slide = as.factor(subgroupsce$Slide))

p_countbygroup = ggplot(count_cell_df, aes(x = totalcount, color = Region)) +
  geom_histogram(aes(y=..density..), fill= "white", alpha=0.5, position="identity") +
  geom_density(alpha = 0.2) +
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  scale_colour_manual(values = region_palette[c(group1, group2)]) +
  xlim(c(0,5000))
```
```{r}
# gene count per cell by donor
p_countbydonor = ggplot(count_cell_df, aes(x = totalcount, color = Donor)) +
  geom_histogram(aes(y=..density..),fill="white", alpha=0.5, position="identity") + 
  geom_density(alpha = 0.2) + 
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  scale_color_manual(values = donors_palette) + 
  xlim(c(0,5000))
```
```{r}
# gene count per cell by slide
p_countbyslide = ggplot(count_cell_df, aes(x = totalcount, color = Slide)) +
  geom_histogram(aes(y=..density..),fill="white", alpha=0.5, position="identity") + 
  geom_density(alpha = 0.2) + 
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlim(c(0,5000))
```
```{r}
# zero proportion per gene
zeroprop_df = rbind(data.frame(zero_prop = rowMeans(assays(subgroupsce)$counts[, subgroupsce$region %in% group1]== 0), genemean = rowMeans(assays(subgroupsce)$counts[, subgroupsce$region %in% group1]), Region = group1),
                data.frame(zero_prop = rowMeans(assays(subgroupsce)$counts[, subgroupsce$region %in% group2]== 0), genemean = rowMeans(assays(subgroupsce)$counts[, subgroupsce$region %in% group2]), Region = group2),
                data.frame(zero_prop = rowMeans(assays(subgroupsce)$counts == 0), genemean = rowMeans(assays(subgroupsce)$counts), Region = "Combined")
                )
zeroprop_df = zeroprop_df[which(zeroprop_df$genemean>0),]
zeroprop_df$Region =factor(zeroprop_df$Region, levels = c(group1, group2, "Combined"))

p_zeroprop = ggplot(zeroprop_df, aes(x = genemean, y = zero_prop)) + 
  geom_point(aes(shape = Region, color = Region), alpha=0.5, size = 1.2) + 
  scale_color_manual(values = c(region_palette[c(group1, group2)],Combined = "darkmagenta")) +
  scale_shape_manual(values = c(16,16,3)) +
  xlab("Gene mean") + 
  ylab("Zero proportion") +
  xlim(c(0,10)) + 
  stat_function(fun = function(x) exp(-x), colour = "black") +
  theme_bw() + theme(panel.grid = element_blank())

```

```{r, fig.width=12, fig.height=8}
ggarrange(p_zeroprop, p_countbygroup, p_countbyslide, p_countbydonor, nrow = 2, ncol = 2)
```

# Difference in library size
```{r}
ks_result = ks.test(subset(count_cell_df, Region == group1, totalcount)$totalcount, subset(count_cell_df, Region == group2, totalcount)$totalcount)
t_test = t.test(subset(count_cell_df, Region == group1, totalcount)$totalcount, subset(count_cell_df, Region == group2, totalcount)$totalcount)
print(ks_result)
print(t_test)
```

# Number of DEGs from each method
```{r}
dflist = list(pois_glmm_df, 
              binom_glmm_df, 
              DESeq2.muscat,
              edgeR.cpm,
              MAST_donor_df, 
              Wilcox_Seurat_df,
              # MMdream,
              MMvst
              # MMpoisson
              )
titlelist = c("Poisson-glmm_UMI", 
              "Binomial-glmm_UMI",
              "pb-DESeq2_UMI", 
              "pb-edgeR_CPM",
              "MAST_CPM",
              "Wilcox_Norm",
              # "MMdream",
              "MMvst_VST"
              # "MMpoisson_UMI"
              )


number_input_hits = data.frame(DEGs = integer(0), Input = integer(0))
pois_glmm_inputgene = pois_glmm_df$genes[!is.na(pois_glmm_df$hits)]
for (df in dflist){
  number_input_hits = rbind(number_input_hits, 
                            c(sum(!is.na(df$hits)), 
                              sum(df$hits, na.rm = T))) 
}
colnames(number_input_hits) = c("Inputs", "DEGs")
number_input_hits$method = titlelist

number_input_hits = melt(number_input_hits, id = "method")
colnames(number_input_hits)[2:3] = c("Gene", "Number_of_genes")
```


```{r}
# pdf("../submission preparation/graph_panels/panel4/input_hits.pdf", width = 8, height = 8)
ggplot(data=number_input_hits, aes(x=factor(method, 
                                            levels = titlelist), y=Number_of_genes, fill = Gene)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label=Number_of_genes), vjust=-0.2, position=position_dodge(0.9), size = 2) +
  theme_bw() + theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  scale_fill_manual(values = c("#BCAAA4", "#5D4037")) +
  xlab("") +
  ylab("Number of genes") +
  ggtitle(paste("DEGs summary"))
# dev.off()

```

# Volcano plot
```{r,fig.width = 10, fig.height=6}
pvalcutoff = 0.05
v = 2
log2FCcutoff = log2(v)

p = list()
for (i in 1:length(dflist)){
  if(i %in% 6){
    xvar = 'avg_log2FC'
    yvar = 'p_val_adj'
  }else{
    xvar = 'log2FC'
    yvar = 'BH'
  }
  if (nrow(dflist[[i]])==0){next}
  p[[i]] = ggplot(na.omit(data.frame(dflist[[i]])), aes_string(x = xvar, y = sprintf("-log10(%s)", yvar), colour = "hits")) +
    geom_point(alpha = 0.5, size = 0.5) +
    labs(color = "hits") +
    geom_hline(yintercept=-log10(pvalcutoff),linetype=2) +
    geom_vline(xintercept=log2FCcutoff,linetype=2) + 
    geom_vline(xintercept=-log2FCcutoff,linetype=2) +
    ggtitle(titlelist[i]) + 
    theme_bw() + 
    theme(panel.grid = element_blank()) + 
    scale_color_manual(values = c("gray", "blue")) +
    labs(color = "DEGs") +
    xlim(c(-3,3)) + 
    ylim(c(0,50)) + 
    xlab("Log2 Fold Change") + 
    ylab("-Log10 (adj.pval)")
}

figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)

# Annotate the figure by adding a common labels
annotate_figure(figure,
                top = paste("Volcano plot"),
                bottom = text_grob(paste( "log2FC cutoff ", paste0("log2(",v,")"),"; adj.pval cutoff", pvalcutoff)
                                   , x = 0.99, just = "right")
)
```


# Histogram of p-value/adj.p-value
```{r, fig.width = 10, fig.height=6}
p = list()
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]])==0){next}
  df = data.frame(genes = dflist[[i]]$genes)
  
  if(i %in% 6){
    df = cbind(df, dflist[[i]][,c("p_val", "p_val_adj")])
  }else{
    df = cbind(df, dflist[[i]][,c("pval", "BH")])
  }
  
  p[[i]] = ggplot(melt(df,id = "genes"), aes(x=value, color=variable)) +
geom_histogram(fill="white", alpha=0.5, position="identity") + theme_bw() + theme(panel.grid = element_blank()) + xlab(NULL) + ylab("Count") + 
coord_cartesian(ylim = c(0, 3000)) +
    ggtitle(titlelist[i]) + scale_colour_discrete(guide = guide_legend(title = NULL),labels = c("p-value", "adj.p-value"))
}
figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)

annotate_figure(figure)
```

# P-Value comparison across different methods
```{r,fig.width = 10, fig.height=6}
value = "pval"
p = list()
  for (i in 2:length(dflist)){
    if (nrow(dflist[[i]])==0){next}
    if (value == "pval"){
      xlimit = c(0,1)
      ylimit = c(0,1)
      if (is.null(dflist[[i]]$pval)){
        x = dflist[[i]]$p_val
      }else
        x = dflist[[i]]$pval
    }
    if (value == "log2FC"){
      xlimit = c(-3,3)
      ylimit = c(-3,3)
      if (is.null(dflist[[i]]$log2FC)){
        x = dflist[[i]]$avg_log2FC
      }else
        x = dflist[[i]]$log2FC
    }
    y = dflist[[1]][match(dflist[[i]]$genes, dflist[[1]]$genes),value]
    p[[i-1]] = ggplot(na.omit(data.frame(x = x, y = y)), aes(x = x, y = y)) +
      geom_point(alpha = 0.5, size = 0.5) + theme_minimal() + xlim(xlimit) + ylim(ylimit)+ 
      xlab(titlelist[i]) + ylab(titlelist[1])
  }
  figure = ggarrange(plotlist = p,nrow = 2, ncol = 4)
  
  # Annotate the figure by adding a common labels
  annotate_figure(figure,
                  top = paste(value, "comparison"))
```

# Log2 fold change comparison across different methods
```{r,fig.width = 10, fig.height=6}
value = "log2FC"
p = list()
  for (i in 2:length(dflist)){
    if (nrow(dflist[[i]])==0){next}
    if (value == "pval"){
      xlimit = c(0,1)
      ylimit = c(0,1)
      if (is.null(dflist[[i]]$pval)){
        x = dflist[[i]]$p_val
      }else
        x = dflist[[i]]$pval
    }
    if (value == "log2FC"){
      xlimit = c(-3,3)
      ylimit = c(-3,3)
      if (is.null(dflist[[i]]$log2FC)){
        x = dflist[[i]]$avg_log2FC
      }else
        x = dflist[[i]]$log2FC
    }
    y = dflist[[1]][match(dflist[[i]]$genes, dflist[[1]]$genes),value]
    p[[i-1]] = ggplot(na.omit(data.frame(x = x, y = y)), aes(x = x, y = y)) +
      geom_point(alpha = 0.5, size = 0.5) + theme_minimal() + xlim(xlimit) + ylim(ylimit)+ 
      xlab(titlelist[i]) + ylab(titlelist[1])
  }
  figure = ggarrange(plotlist = p,nrow = 2, ncol = 4)
  
  # Annotate the figure by adding a common labels
  annotate_figure(figure,
                  top = paste(value, "comparison"))
```

# Violin plot of log2mean of DEGs
```{r, fig.width=7}
p = list()
df = data.frame()
clustername = pair
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]])==0){next}
  if (sum(dflist[[i]]$hits, na.rm = T)==0){next}
  hits_gene = dflist[[i]]$genes[which(dflist[[i]]$hits)]
  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:2){
    cellind = subgroupsce$region %in% clustername[j]
    df = rbind(df, data.frame(Log2mean = log2(fun(assays(subgroupsce)$counts[hits_gene,cellind])), Group = clustername[j], method = titlelist[i]))
  }
}
```


```{r, fig.width=7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=Log2mean, fill = factor(Group))) +
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylim(c(-3,4)) + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  labs(fill = "Region") +
  scale_fill_manual(values = region_palette[c(group1, group2)])
```

# Violin plot of gene expression frequency of DEGs
```{r, fig.width = 7}
p = list()
df = data.frame()
clusterlist = list(unique(sce_human_spine_filtered$region), group1, group2)
clustername = c("All region", group1, group2)
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]])==0){next}
  if (sum(dflist[[i]]$hits, na.rm = T)==0){next}
  hits_gene = dflist[[i]]$genes[which(dflist[[i]]$hits)]

  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:3){
    cellind = sce_human_spine_filtered$region %in% clusterlist[[j]]
    df = rbind(df, data.frame(freq = fun(assays(sce_human_spine_filtered)$counts[hits_gene,cellind]>0), cluster = clustername[j], method = titlelist[i]))
  }
}
levels(df$cluster) = clustername
```



```{r, fig.width = 7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=freq, fill = factor(cluster, levels =  levels(cluster)))) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylab("Gene expression frequency") + 
  ylim(c(0,1)) + 
  xlab("") + 
  scale_fill_manual(values = c("All region" = "gray", region_palette[c(group1, group2)]), breaks = levels(df$cluster)) +
  labs(fill = "Region") +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) 
```

# Upset plot
```{r}
library(UpSetR)
all_genes = c()
for(df in dflist){
  all_genes = c(all_genes, df$genes[which(df$hits)])
}

all_genes = unique(all_genes)

# Create a binary matrix where rows are genes and columns represent methods
upset_data = data.frame(gene = all_genes)
for (i in seq_along(dflist)){
  title = titlelist[i]
  upset_data[[title]] = as.numeric(all_genes %in% dflist[[i]]$genes[which(dflist[[i]]$hits)])
}

# Generate the upset plot
upset(upset_data[,-1], 
      sets = rev(titlelist), 
      keep.order = TRUE, 
      order.by = "freq")

```

# Heatmap of top DEGs

## Poisson-glmm DEGs

```{r}
sort_log2FC = sort(abs(pois_glmm_df$log2FC[pois_glmm_df$hits]), index.return = T, decreasing = T)
hitgenes = pois_glmm_df$genes[which(pois_glmm_df$hits)][sort_log2FC$ix]
mat = subgroupsce@assays@data$counts[hitgenes,]
#row annotation
annotation_log2FC = data.frame(abs_log2FC = abs(pois_glmm_df$log2FC[match(hitgenes,pois_glmm_df$genes)]))
rownames(annotation_log2FC) = rownames(mat)
#col annotation
annotation_df = data.frame(Donor = subgroupsce$patient_ID, Region = subgroupsce$region)
rownames(annotation_df) = colnames(mat)
annotation_df = annotation_df[with(annotation_df, order(Region, Donor)), ]
# cell level
pheatmap(mat[,rownames(annotation_df)],
         main = paste0("Poisson-glmm DEGs\nin UMI counts (", nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```

## Additional DEGs from other methods

### Wilcox
```{r}
sort_log2FC = sort(abs(Wilcox_Seurat_df$avg_log2FC[Wilcox_Seurat_df$hits]), index.return = T, decreasing = T)
hitgenes_wilcox = Wilcox_Seurat_df$genes[which(Wilcox_Seurat_df$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_wilcox[!hitgenes_wilcox%in%hitgenes])
mat = subgroupsce@assays@data$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Wilcox DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```

### pb-DESeq2
```{r}
sort_log2FC = sort(abs(DESeq2.muscat$log2FC[DESeq2.muscat$hits]), index.return = T, decreasing = T)
hitgenes_deseq2 = DESeq2.muscat$genes[which(DESeq2.muscat$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_deseq2[!hitgenes_deseq2%in%hitgenes][1:600])
mat = subgroupsce@assays@data$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("pb-DESeq2 DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```
```{r eval=FALSE}
pdf("~/Google Drive/My Drive/spatial/10X/graph_panels/panel7/heatmap_deseq2.pdf", width = 5, height = 5)
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("pb-DESeq2 DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
dev.off()
```

# MA plot 
```{r, fig.width = 10, fig.height=5}
xvar = subgroupsce@metadata$log2mean
xlimits = c(-6,6)
xLabel = "Log2 mean"
p = list()
  for (i in 1:length(dflist)){
    if(is.null(dflist[[i]]$genes)){
      x = xvar[match(rownames(dflist[[i]]), names(xvar))]
    }else{
      x = xvar[match(dflist[[i]]$genes,names(xvar))]
    }
    if(is.null(dflist[[i]]$log2FC)){
      y = dflist[[i]]$avg_log2FC
    }else{
      y = dflist[[i]]$log2FC
    }
    p[[i]] = ggplot(na.omit(data.frame(x = x, y = y, DEGs = dflist[[i]]$hits)), aes(x = x, y = y, colour = DEGs)) +
      geom_point(alpha = 0.5, size = 0.5) +
      geom_hline(yintercept=log2(1.5),linetype=2) +
      geom_hline(yintercept=-log2(1.5),linetype=2) +
      ggtitle(titlelist[i]) + theme_minimal() + xlim(xlimits) + ylim(c(-4,4)) + 
      scale_color_manual(values = c("gray", "blue")) +
      xlab(xLabel) + ylab("Log2 Fold Change") + theme(legend.position = "bottom")
  }
  fig_FC_mean = ggarrange(plotlist = p,common.legend = TRUE, legend = "right", ncol = 4, nrow = 2)
  annotate_figure(fig_FC_mean, top = paste("Log2FC vs. Log2 mean "))
```

# Enrichment analysis

## GO object
```{r, fig.width = 6, fig.height=8}
original_gene_list = pois_glmm_df$log2FC
names(original_gene_list) = pois_glmm_df$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
genes = pois_glmm_df$genes[which(pois_glmm_df$hits)]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)

```
```{r, fig.width = 8, fig.height=6}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", color.params = list(foldChange = gene_list))
```


## enrichKEGG object
```{r, fig.width = 10, fig.height=5}
# Convert gene IDs for enrichKEGG function
# We will lose some genes here because not all IDs will be converted
ids = bitr(names(original_gene_list), fromType = "SYMBOL", toType = "ENTREZID", OrgDb="org.Hs.eg.db") 
dedup_ids = ids[!duplicated(ids[c("SYMBOL")]),]

# Create a vector of the gene universe
match_logic = pois_glmm_df$genes %in% dedup_ids$SYMBOL
kegg_gene_list <- pois_glmm_df$log2FC[match_logic]
names(kegg_gene_list) <- dedup_ids$ENTREZID
kegg_gene_list<-na.omit(kegg_gene_list)
kegg_gene_list = sort(kegg_gene_list, decreasing = TRUE)
# hits
kegg_genes = na.omit(dedup_ids$ENTREZID[pois_glmm_df$hits[match_logic]])
# create enrichKEGG object
kk = enrichKEGG(gene=kegg_genes,
                universe=names(kegg_gene_list),
                organism="hsa", pvalueCutoff = 0.05, keyType = "ncbi-geneid")
```
```{r, fig.width = 8, fig.height=4}
dotplot(kk, 
        showCategory = 10, 
        title = "Enriched Pathways",
        font.size = 8)
```

```{r, fig.width = 10, fig.height=5}
cnetplot(kk, categorySize="pvalue", foldChange=kegg_gene_list)
```
