---
title: "P-values and false discovery rates assessed under the null setting using permutation analysis"
author: "Chih-Hsuan Wu"
output: html_document
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
source("./code/hippo_all_proportion.R")
source("./code/DE_methods.R")
source("./code/functions_in_rmd.R")

library(ggplot2)
library(dplyr)
library(ggpubr)
library(SingleCellExperiment)
library(pheatmap)
library(gridExtra)
library(reshape)
library(Seurat)
library(SeuratObject)
library(muscat)
library(MAST)
library(tidyr)
```

```{r, eval = FALSE}
# Load cluster 2 and 13 in case study 1
load("./data/10X_inputdata.RData")
load("./data/10X_inputdata_integrated.RData")
load("./data/vstcounts.RData")
```
```{r, eval = FALSE}
cellgroup = 2
subgroupsce = inputData[, inputData$hippo_cluster%in% cellgroup]

# Run simulation on one cluster in case sdudy 1
set.seed(20231129)
sim_time = 20
pval_result = setNames(data.frame(matrix(ncol = 8, nrow = sim_time)), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
FD_result = setNames(data.frame(matrix(ncol = 8, nrow = sim_time)), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
poisglmm_pval = c()
binomglmm_pval = c()
pb_DESeq2_pval = c()
pb_edgeR_pval = c()
MAST_pval = c()
Wilcox_pval = c()
MMvst_pval = c()
MMpoisson_pval = c()
labels = matrix(nrow = sim_time, ncol = length(subgroupsce$donor))

# generate permuted labels
for(i in 1:sim_time){
  labels[i,] = rep("stim",length(subgroupsce$donor))
  labels[i,sample(1:length(subgroupsce$donor), round(length(subgroupsce$donor)/2))] = "ctrl"
}
  subgroupsce$group_id = as.factor(labels[1,])
  subgroupsce$cluster_id = "void"
  subgroupsce$sample_id = as.factor(paste0(subgroupsce$group_id, subgroupsce$donor))
  subgroupsce = prepSCE(subgroupsce,
    kid = "cluster_id", # subpopulation assignments
    gid = "group_id",  # group IDs (ctrl/stim)
    sid = "sample_id",   # sample IDs (ctrl/stim 1234)
    drop = FALSE)  # keep all other colData columns
  
for(i in 1:sim_time){
  # assign labels and adjust metadata in sce
  subgroupsce$group_id = as.factor(labels[i,])
  subgroupsce$sample_id = as.factor(paste0(subgroupsce$group_id, subgroupsce$donor))
  subgroupsce = prepSCE(subgroupsce,
    kid = "cluster_id", # subpopulation assignments
    gid = "group_id",  # group IDs (ctrl/stim)
    sid = "sample_id",   # sample IDs (ctrl/stim 1234)
    drop = FALSE)  # keep all other colData columns
  genemean1 = rowMeans(counts(subgroupsce)[,subgroupsce$group_id == "ctrl"])
  genemean2 = rowMeans(counts(subgroupsce)[,subgroupsce$group_id == "stim"])
  subgroupsce@metadata$log2mean = log2(genemean1*genemean2)/2
  subgroupsce@metadata$log2meandiff = log2(abs(genemean1-genemean2))

  
  subgroupsce_vst = subgroupsce[rownames(subgroupsce)%in% rownames(vstcounts),]
assays(subgroupsce_vst)$vstresiduals = vstcounts[, colnames(vstcounts)%in% colnames(subgroupsce_vst)]
subgroupsce_integrated = inputData_integrated[,colnames(subgroupsce)]

  
  
  
  # poisson_glmm
  pois_glmm_df = poisson_glmm_DE(subgroupsce, subgroupsce$group_id, subgroupsce$donor)
  pois_glmm_df$hits = identifyDEGs(pois_glmm_df$BH, pois_glmm_df$log2FC, subgroupsce@metadata$log2mean, subgroupsce@metadata$log2meandiff, newcriteria = T, log2FCcutoff = 1)
  input_genes = pois_glmm_df$status=="done"
  pval_result[i,"Poisson-glmm"] = mean(pois_glmm_df$pval<0.05, na.rm = T)
  FD_result[i,"Poisson-glmm"] = mean(pois_glmm_df$hits==T, na.rm = T)
  poisglmm_pval = c(poisglmm_pval, na.omit(pois_glmm_df$pval))
  
  # binom_glmm
  binom_glmm_df = binomial_glmm_DE(subgroupsce[input_genes,], subgroupsce$group_id, subgroupsce$donor)
  binom_glmm_df$hits = identifyDEGs(binom_glmm_df$BH, binom_glmm_df$log2FC, subgroupsce@metadata$log2mean, subgroupsce@metadata$log2meandiff, newcriteria = T, log2FCcutoff = 1)
  pval_result[i,"Binomial-glmm"] = mean(binom_glmm_df$pval<0.05, na.rm = T)
  FD_result[i,"Binomial-glmm"] = mean(binom_glmm_df$hits==T, na.rm = T)
  binomglmm_pval = c(binomglmm_pval, na.omit(binom_glmm_df$pval))
  
  # MAST
  MAST_df = MAST_DE(subgroupsce[input_genes,], subgroupsce$group_id, subgroupsce$donor)
  MAST_df$hits = identifyDEGs(MAST_df$BH, MAST_df$log2FC, log2FCcutoff = 1)
  pval_result[i,"MAST"] = mean(MAST_df$pval<0.05, na.rm = T)
  FD_result[i,"MAST"] = mean(MAST_df$hits==T, na.rm = T)
  MAST_pval = c(MAST_pval, na.omit(MAST_df$pval))
  # Seurat
  Seurat_subgroup_int = CreateSeuratObject(counts(subgroupsce_integrated), project = as.character(subgroupsce$group_id), assay = "RNA", meta.data = data.frame(colData(subgroupsce)))
  
  # Wilcox
  Wilcox_Seurat_df = FindMarkers(Seurat_subgroup_int, ident.1 = as.character("ctrl"), ident.2 = as.character("stim"), test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
  Wilcox_Seurat_df$hits = identifyDEGs(Wilcox_Seurat_df$p_val_adj, Wilcox_Seurat_df$avg_log2FC, log2FCcutoff = 1)
  pval_result[i,"Wilcox"] = mean(Wilcox_Seurat_df$p_val<0.05, na.rm = T)
  FD_result[i,"Wilcox"] = mean(Wilcox_Seurat_df$hits==T, na.rm = T)
  Wilcox_pval = c(Wilcox_pval, na.omit(Wilcox_Seurat_df$p_val))
  
    # DESeq2.muscat
  pb = aggregateData(subgroupsce[input_genes,],
    assay = "counts", fun = "sum",
    by = c("cluster_id", "sample_id"))
  res = pbDS(pb, method = "DESeq2", verbose = F)
  DESeq2.muscat = data.frame(res$table[[1]][[1]])[,c(1,4,7,8)]
  DESeq2.muscat$logFC = log2(exp(DESeq2.muscat$logFC))
  colnames(DESeq2.muscat) = c("genes", "log2FC", "pval", "BH")
  DESeq2.muscat$hits = identifyDEGs(DESeq2.muscat$BH, DESeq2.muscat$log2FC, log2FCcutoff = 1)
  pval_result[i,"pb-DESeq2"] = mean(DESeq2.muscat$pval<0.05, na.rm = T)
  FD_result[i,"pb-DESeq2"] = mean(DESeq2.muscat$hits==T, na.rm = T)
  pb_DESeq2_pval = c(pb_DESeq2_pval, na.omit(DESeq2.muscat$pval))
  
  # edgeR.sum.cpm
  pb = aggregateData(subgroupsce[input_genes,],
    assay = "cpm", fun = "sum",
    by = c("cluster_id", "sample_id"))
  res = pbDS(pb, method = "edgeR", verbose = FALSE)
  edgeR.cpm = data.frame(res$table[[1]][[1]])[,c(1,3,6,7)]
  edgeR.cpm$logFC = log2(exp(edgeR.cpm$logFC))
  colnames(edgeR.cpm)[1:4] = c("genes", "log2FC", "pval", "BH")
  edgeR.cpm$hits = identifyDEGs(edgeR.cpm$BH, edgeR.cpm$log2FC, log2FCcutoff = 1)
  pval_result[i,"pb-edgeR"] = mean(edgeR.cpm$pval<0.05, na.rm = T)
  FD_result[i,"pb-edgeR"] = mean(edgeR.cpm$hits==T, na.rm = T)
  pb_edgeR_pval = c(pb_edgeR_pval, na.omit(edgeR.cpm$pval))
  
    # MM-vst
  res = mmDS(subgroupsce_vst, method = "vst", vst = "sctransform")
  MMvst = data.frame(res[[1]])[,c(1,3,6,7)]
  MMvst$logFC = log2(exp(MMvst$beta))
  colnames(MMvst)[1:4] = c("genes", "log2FC", "pval", "BH")
  MMvst$hits = identifyDEGs(MMvst$BH, MMvst$log2FC, log2FCcutoff = 1)
  pval_result[i,"MMvst"] = mean(MMvst$pval<0.05, na.rm = T)
  FD_result[i,"MMvst"] = mean(MMvst$hits==T, na.rm = T)
  MMvst_pval = c(MMvst_pval, na.omit(MMvst$pval))
  
  # MM-poisson
  res = mmDS(subgroupsce[input_genes,], method = "poisson")
  MMpoisson = data.frame(res[[1]])[,c(1,3,6,7)]
  MMpoisson$beta = log2(exp(MMpoisson$beta))
  colnames(MMpoisson)[1:4] = c("genes", "log2FC", "pval", "BH")
  MMpoisson$hits = identifyDEGs(MMpoisson$BH,
                                            MMpoisson$log2FC, log2FCcutoff = 1)
  pval_result[i,"MMpoisson"] = mean(MMpoisson$pval<0.05, na.rm = T)
  FD_result[i,"MMpoisson"] = mean(MMpoisson$hits==T, na.rm = T)
  MMpoisson_pval = c(MMpoisson_pval, na.omit(MMpoisson$pval))

}

# aggregate p-value results
pval_list = list(poisglmm_pval, binomglmm_pval, pb_DESeq2_pval, pb_edgeR_pval, MAST_pval, Wilcox_pval, MMvst_pval, MMpoisson_pval)
max_length = max(sapply(pval_list,length))
pval_list = sapply(pval_list, function(x){
    c(x, rep(NA, max_length - length(x)))
})
pval_df = setNames(as.data.frame(pval_list), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
save(FD_result, pval_result, pval_df, file = paste0("./data/permutation",cellgroup,".RData"))
```
```{r, eval = FALSE}
# Load Bcells in Kang dataset
load("./data/data2sce.Rdata")
subgroupsce <<- sce[, sce$cluster_id%in% "B cells" & sce$group_id %in% "ctrl"]
subgroupsce$ind = paste0("D",subgroupsce$ind)
# only use ctrl group
rm(sce)

# Run simulation on B cells in Kang dataset
set.seed(20231129)
sim_time = 20
pval_result = setNames(data.frame(matrix(ncol = 8, nrow = sim_time)), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
FD_result = setNames(data.frame(matrix(ncol = 8, nrow = sim_time)), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
poisglmm_pval = c()
binomglmm_pval = c()
pb_DESeq2_pval = c()
pb_edgeR_pval = c()
MAST_pval = c()
Wilcox_pval = c()
MMvst_pval = c()
MMpoisson_pval = c()
labels = matrix(nrow = sim_time, ncol = length(subgroupsce$group_id))

# generate permuted labels
for(i in 1:sim_time){
  labels[i,] = rep("stim",length(subgroupsce$group_id))
  labels[i,sample(1:length(subgroupsce$group_id), length(subgroupsce$group_id)/2)] = "ctrl"
}

for(i in 1:sim_time){
  # assign labels and adjust metadata in sce
  subgroupsce$group_id = as.factor(labels[i,])
  subgroupsce$sample_id = as.factor(paste0(subgroupsce$group_id, subgroupsce$ind))
  subgroupsce@metadata$experiment_info$sample_id = unique(subgroupsce$sample_id)
  subgroupsce@metadata$experiment_info$group_id = as.factor(substr(unique(subgroupsce$sample_id), 1,4))
  genemean1 = rowMeans(counts(subgroupsce)[,subgroupsce$group_id=="ctrl"])
  genemean2 = rowMeans(subgroupsce@assays@data$counts[,subgroupsce$group_id=="stim"])
  subgroupsce@metadata$log2mean = log2(genemean1*genemean2)/2
  subgroupsce@metadata$log2meandiff = log2(abs(genemean1-genemean2))

  # poisson_glmm
  Kang_pois_glmm_df = hippo_poisson_glmm_DE(subgroupsce, subgroupsce$group_id, subgroupsce$ind)
  Kang_pois_glmm_df$hits = identifyDEGs(Kang_pois_glmm_df$BH, Kang_pois_glmm_df$log2FC, subgroupsce@metadata$log2mean, subgroupsce@metadata$log2meandiff, newcriteria = T, log2FCcutoff = 1)
  input_genes = Kang_pois_glmm_df$status=="done"
  pval_result[i,"poisglmm"] = mean(Kang_pois_glmm_df$pval<0.05, na.rm = T)
  FD_result[i,"poisglmm"] = mean(Kang_pois_glmm_df$hits==T, na.rm = T)
  poisglmm_pval = c(poisglmm_pval, na.omit(Kang_pois_glmm_df$pval))
  
  # binom_glmm
  Kang_binom_glmm_df = hippo_binomial_glmm_DE(subgroupsce[input_genes,], subgroupsce$group_id, subgroupsce$ind)
  Kang_binom_glmm_df$hits = identifyDEGs(Kang_binom_glmm_df$BH, Kang_binom_glmm_df$log2FC, subgroupsce@metadata$log2mean, subgroupsce@metadata$log2meandiff, newcriteria = T, log2FCcutoff = 1)
  pval_result[i,"binomglmm"] = mean(Kang_binom_glmm_df$pval<0.05, na.rm = T)
  FD_result[i,"binomglmm"] = mean(Kang_binom_glmm_df$hits==T, na.rm = T)
  binomglmm_pval = c(binomglmm_pval, na.omit(Kang_binom_glmm_df$pval))
  
  # MAST
  Kang_MAST_df = MAST_DE(subgroupsce[input_genes,], subgroupsce$group_id, subgroupsce$ind)
  Kang_MAST_df$hits = identifyDEGs(Kang_MAST_df$BH, Kang_MAST_df$log2FC, log2FCcutoff = 1)
  pval_result[i,"MAST"] = mean(Kang_MAST_df$pval<0.05, na.rm = T)
  FD_result[i,"MAST"] = mean(Kang_MAST_df$hits==T, na.rm = T)
  MAST_pval = c(MAST_pval, na.omit(Kang_MAST_df$pval))
  # Seurat
  Seurat_subgroup_lognorm = CreateSeuratObject(subgroupsce@assays@data$logcounts[input_genes,], project = as.character(subgroupsce$group_id), assay = "RNA", meta.data = data.frame(colData(subgroupsce)))
  
  # Wilcox
  Kang_Wilcox_Seurat_df = FindMarkers(Seurat_subgroup_lognorm, ident.1 = as.character("ctrl"), ident.2 = as.character("stim"), test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
  Kang_Wilcox_Seurat_df$hits = identifyDEGs(Kang_Wilcox_Seurat_df$p_val_adj, Kang_Wilcox_Seurat_df$avg_log2FC, log2FCcutoff = 1)
  pval_result[i,"Wilcox"] = mean(Kang_Wilcox_Seurat_df$p_val<0.05, na.rm = T)
  FD_result[i,"Wilcox"] = mean(Kang_Wilcox_Seurat_df$hits==T, na.rm = T)
  Wilcox_pval = c(Wilcox_pval, na.omit(Kang_Wilcox_Seurat_df$p_val))
  
    # DESeq2.muscat
  pb = aggregateData(subgroupsce[input_genes,],
    assay = "counts", fun = "sum",
    by = c("cluster_id", "sample_id"))
  res = pbDS(pb, method = "DESeq2", verbose = F)
  Kang_DESeq2.muscat = data.frame(res$table[[1]][[1]])[,c(1,4,7,8)]
  Kang_DESeq2.muscat$logFC = log2(exp(Kang_DESeq2.muscat$logFC))
  colnames(Kang_DESeq2.muscat) = c("genes", "log2FC", "pval", "BH")
  Kang_DESeq2.muscat$hits = identifyDEGs(Kang_DESeq2.muscat$BH, Kang_DESeq2.muscat$log2FC, log2FCcutoff = 1)
  pval_result[i,"pb_DESeq2"] = mean(Kang_DESeq2.muscat$pval<0.05, na.rm = T)
  FD_result[i,"pb_DESeq2"] = mean(Kang_DESeq2.muscat$hits==T, na.rm = T)
  pb_DESeq2_pval = c(pb_DESeq2_pval, na.omit(Kang_DESeq2.muscat$pval))
  
  # edgeR.sum.cpm
  assays(subgroupsce)$cpm = edgeR::cpm(subgroupsce@assays@data$counts)
  pb = aggregateData(subgroupsce[input_genes],
    assay = "cpm", fun = "sum",
    by = c("cluster_id", "sample_id"))
  res = pbDS(pb, method = "edgeR", verbose = FALSE)
  Kang_edgeR.cpm = data.frame(res$table[[1]][[1]])[,c(1,3,6,7)]
  Kang_edgeR.cpm$logFC = log2(exp(Kang_edgeR.cpm$logFC))
  colnames(Kang_edgeR.cpm)[1:4] = c("genes", "log2FC", "pval", "BH")
  Kang_edgeR.cpm$hits = identifyDEGs(Kang_edgeR.cpm$BH, Kang_edgeR.cpm$log2FC, log2FCcutoff = 1)
  pval_result[i,"pb_edgeR"] = mean(Kang_edgeR.cpm$pval<0.05, na.rm = T)
  FD_result[i,"pb_edgeR"] = mean(Kang_edgeR.cpm$hits==T, na.rm = T)
  pb_edgeR_pval = c(pb_edgeR_pval, na.omit(Kang_edgeR.cpm$pval))

  
  # MM-vst
  res = mmDS(subgroupsce[input_genes], method = "vst", vst = "sctransform")
  Kang_MMvst = data.frame(res[[1]])[,c(1,3,6,7)]
  Kang_MMvst$logFC = log2(exp(Kang_MMvst$beta))
  colnames(Kang_MMvst)[1:4] = c("genes", "log2FC", "pval", "BH")
  Kang_MMvst$hits = identifyDEGs(Kang_MMvst$BH, Kang_MMvst$log2FC, log2FCcutoff = 1)
  pval_result[i,"MMvst"] = mean(Kang_MMvst$pval<0.05, na.rm = T)
  FD_result[i,"MMvst"] = mean(Kang_MMvst$hits==T, na.rm = T)
  MMvst_pval = c(MMvst_pval, na.omit(Kang_MMvst$pval))
  
  # MM-poisson
  res = mmDS(subgroupsce[input_genes,], method = "poisson")
  Kang_MMpoisson = data.frame(res[[1]])[,c(1,3,6,7)]
  Kang_MMpoisson$beta = log2(exp(Kang_MMpoisson$beta))
  colnames(Kang_MMpoisson)[1:4] = c("genes", "log2FC", "pval", "BH")
  Kang_MMpoisson$hits = identifyDEGs(Kang_MMpoisson$BH,
                                            Kang_MMpoisson$log2FC, log2FCcutoff = 1)
  pval_result[i,"MMpoisson"] = mean(Kang_MMpoisson$pval<0.05, na.rm = T)
  FD_result[i,"MMpoisson"] = mean(Kang_MMpoisson$hits==T, na.rm = T)
  MMpoisson_pval = c(MMpoisson_pval, na.omit(Kang_MMpoisson$pval))
}

# aggregate p-value results
pval_list = list(poisglmm_pval, binomglmm_pval, pb_DESeq2_pval, pb_edgeR_pval, MAST_pval, Wilcox_pval, MMvst_pval, MMpoisson_pval)
max_length = max(sapply(pval_list,length))
pval_list = sapply(pval_list, function(x){
    c(x, rep(NA, max_length - length(x)))
})
pval_df = setNames(as.data.frame(pval_list), c("Poisson-glmm", "Binomial-glmm", "pb-DESeq2", "pb-edgeR", "MAST", "Wilcox", "MMvst", "MMpoisson"))
save(FD_result, pval_result, pval_df, file = "./data/permutation.RData")
```
To examine the p-value calibration in real data, we did a permutation on group-of-interest within a null dataset. The cells in the null dataset were randomly assigned to controlled or stimulated group. We then computed p-values of each gene with different methods. The gene set was restricted to the input genes of Poisson-glmm, and the threshold of Wilcox method was relaxed to prevent filtering out genes. The procedure was repeated 20 times. Each time the proportion of p-value smaller than 0.05 was computed, so as the false discovery DEGs. 

From the violin plot shown below, LEMUR glmm methods and Wilcox method are consistently well-calibrated among different choices of null datasets. However, pseudo-bulk methods, and mixed models from Muscat are too conservative. Their overall proportion is way less than 0.05. The performance of MAST is not consistent among three datasets, which is conservative in B cells but not in case study 1.  The histograms of all p-values in these 20 runs are flat for LEMUR glmm methods and Wilcox method, which satisfy the null setting. However, the p-values of the other methods are overestimated, resulting conservative results. Note that even though Wilcox performs well in the permutation analysis, it is not powerful to detect real DEGs. With either current criteria or our new criteria to determine DEGs, every method detects at most one false discovery each run.

# Different choices of null sets

## Cluster 2 in case study 1
```{r}
load("./data/permutation2.RData")
pval_violin = ggplot(gather(pval_result), aes(value, key)) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value") + 
  ylab("") +
  xlim(c(0,0.08))+
  geom_vline(xintercept  = 0.05, linetype = "dashed", color = "red") +
  scale_y_discrete(limits = rev(unique(gather(pval_result)$key)))
pval_violin
```
```{r}
ggplot(gather(pval_df), aes(value)) +
  geom_histogram() + 
  facet_wrap(~factor(key, unique(gather(pval_df)$key)))+
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value")
```

## Cluster 13 in case study 1
```{r}
load("./data/permutation13.RData")
pval_violin = ggplot(gather(pval_result), aes(value, key)) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value") + 
  ylab("") +
  xlim(c(0,0.08))+
  geom_vline(xintercept  = 0.05, linetype = "dashed", color = "red") +
  scale_y_discrete(limits = rev(unique(gather(pval_result)$key)))
pval_violin
```

```{r}
ggplot(gather(pval_df), aes(value)) +
  geom_histogram() + 
  facet_wrap(~factor(key, unique(gather(pval_df)$key)))+
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value")
```



## The control group of B cells in case study 2
```{r}
load("./data/permutation.RData")
pval_violin = ggplot(gather(pval_result), aes(value, key)) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value") + 
  ylab("") +
  xlim(c(0,0.08))+
  geom_vline(xintercept  = 0.05, linetype = "dashed", color = "red") +
  scale_y_discrete(limits = rev(unique(gather(pval_result)$key)))
pval_violin
```


```{r, eval = FALSE}
pdf("./../graph_panels/panel7/permutation_violin.pdf", width = 5, height = 3)
pval_violin
dev.off()
```


```{r}
ggplot(gather(pval_df), aes(value)) +
  geom_histogram() + 
  facet_wrap(~factor(key, unique(gather(pval_df)$key)))+
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  xlab("p-value")
```

