---
title: "Group12_19"
author: "Chih-Hsuan Wu"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
source("./code/hippo_all_proportion.R")
source("./code/DE_methods.R")
source("./code/functions_in_rmd.R")
load("./data/10X_inputdata.RData")
load("./data/10X_inputdata_integrated.RData")
load("./data/10X_DEresult_update.RData")
load("./data/vstcounts.RData")
library(ggplot2)
library(dplyr)
library(ggpubr)
library(SingleCellExperiment)
library(pheatmap)
library(gridExtra)
library(reshape)
library(clusterProfiler)
library(enrichplot)
library("org.Hs.eg.db")
library(pathview)
assign_subgroup(12,19)
Wilcox_Seurat_integrated_df[[key]]$avg_log2FC = - Wilcox_Seurat_integrated_df[[key]]$avg_log2FC
```
```{r}
combined_palette = c("#E91E63", "#00BCD4")
hippo_palette = c(
  "#A0522D", "#6A00FF", "#A4C400", "#6D8764", "#F472D0",
  "#76608A", "#A20025", "#00ABA9", "#F0A30A", "#008A00",
  "#647687", "#1BA1E2", "#FA6800", "#825A2C", "#0050EF",
  "#E3C800", "#E51400", "#D80073", "#60A917", "#AA00FF")
donors_palette = c("#FDA4AF", "#FCD34D", "#BEF264", "#67E8F9", "#A5B4FC")
annotation_colors = list(Group = c(setNames(hippo_palette[c(group1, group2)], c(group1, group2))), Donor = setNames(donors_palette, paste0("D", 3:7)))
```

# Data summary

```{r}
df = data.frame(Group = subgroupsce$hippo_cluster, Donor = subgroupsce$donor, Count = 1)
df = aggregate(df$Count, list(df$Donor, df$Group), FUN = sum)
colnames(df) = c("Donor", "Group", "Count")
df = df[nrow(df):1,]
df$cumcount= c(t(aggregate(df$Count, list(rev(df$Group)), FUN = cumsum)[,-1]))
p_donor = ggplot(df, aes(x = Group, y = Count, fill = Donor)) + 
  geom_bar(stat = "identity") + 
  xlab("Group") + 
  ylab("Number of cells") + 
  scale_fill_manual(values = donors_palette) + 
  geom_text(aes(label=Count, y = cumcount), vjust = -0.2, size = 2) + 
  theme_bw() + 
  theme(panel.grid = element_blank())
```

```{r}
composition = data.frame(
  zeromean = sum(pois_glmm_df[[key]]$status=="zero mean"),
  lowlyexpressed = sum(pois_glmm_df[[key]]$status=="lowly expressed"), 
  valid = sum(pois_glmm_df[[key]]$status=="done"))
composition = melt(composition)

p_composition = ggplot(composition, aes(x = "", y = value, fill = variable)) + 
  geom_bar(stat = "identity") + 
  xlab("") + 
  ylab("number of genes") + 
  scale_fill_discrete(name = "", labels=c("zero mean", "freq < 5%", "freq > 5%")) +
  geom_text(aes(label=value), position = position_stack(vjust = 0.5), size = 3) + 
  theme_bw() + theme(panel.grid = element_blank()) + 
  ggtitle(paste("gene composition of ", key))
```

```{r}
# gene count per cell by group

count_cell_df = data.frame(totalcount = apply(assays(subgroupsce)$counts, 2, sum), Group = sapply(cellgroup, FUN = function(x)ifelse(x == 1, paste(group1, collapse = "_"), paste(group2, collapse = "_"))), Donor = subgroupsce$donor)
count_cell_df$Group = factor(count_cell_df$Group, levels = c(paste(group1, collapse = "_"), paste(group2, collapse = "_")))

p_countbygroup = ggplot(count_cell_df, aes(x = totalcount, color = Group)) +
  geom_histogram(aes(y=..density..), fill= "white", alpha=0.5, position="identity") +
  geom_density(alpha = 0.2) +
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  scale_colour_manual(values = hippo_palette[c(group1, group2)]) +
  xlim(c(0,5000))
```
```{r}
# gene count per cell by donor
p_countbydonor = ggplot(count_cell_df, aes(x = totalcount, color = Donor)) +
  geom_histogram(aes(y=..density..),fill="white", alpha=0.5, position="identity") + 
  geom_density(alpha = 0.2) + 
  xlab("Library size") + 
  ylab("Density") +
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  scale_color_manual(values = donors_palette) + 
  xlim(c(0,5000))
```
```{r}
# zero proportion per gene
zeroprop_df = rbind(data.frame(zero_prop = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% group1]== 0), genemean = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% group1]), Group = paste(group1, collapse = "_")),
                data.frame(zero_prop = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% group2]== 0), genemean = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% group2]), Group = paste(group2, collapse = "_")),
                data.frame(zero_prop = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% c(group1,group2)]== 0), genemean = rowMeans(assays(inputData)$counts[, inputData$hippo_cluster %in% c(group1,group2)]), Group = key)
                )
zeroprop_df = zeroprop_df[which(zeroprop_df$genemean>0),]
zeroprop_df$Group = factor(zeroprop_df$Group, levels = c(paste(group1, collapse = "_"), paste(group2, collapse = "_"), key))
p_zeroprop = ggplot(zeroprop_df, aes(x = genemean, y = zero_prop, Group = Group)) + 
  geom_point(aes(shape = Group, color = Group), alpha=0.5, size = 1.2) + 
  xlab("Gene mean") + 
  ylab("Zero proportion") +
  scale_color_manual(name = "Group", values = c(hippo_palette[c(group1, group2)],"darkmagenta")) + 
  scale_shape_manual(values = c(16,16,3)) +
  xlim(c(0,10)) + 
  stat_function(fun = function(x) exp(-x), colour = "black") +
  theme_bw() + theme(panel.grid = element_blank())

```

```{r, fig.width=8}
ggarrange(p_donor, p_zeroprop, p_countbygroup, p_countbydonor, nrow = 2, ncol = 2)
```

# Mean difference in raw data/normalized data

```{r}
subset_ind = inputData$hippo_cluster%in% c(group1,group2)
raw_mean_DE = simple_mean_DE(log2(assays(subgroupsce)$counts+1), cellgroup == 1, cellgroup == 2)
integrated_mean_DE = simple_mean_DE(assays(subgroupsce_integrated)$counts, cellgroup == 1, cellgroup == 2)
vst_mean_DE = simple_mean_DE(vstcounts[,subset_ind], cellgroup == 1, cellgroup == 2)
cpm_mean_DE = simple_mean_DE(log2(assays(subgroupsce)$cpm+1), cellgroup == 1, cellgroup == 2)
```
```{r, fig.width=12, fig.height=4}
p_ttest_raw_integrated = ggplot(data = abs(data.frame(raw_t = raw_mean_DE[intersect(rownames(raw_mean_DE), rownames(integrated_mean_DE)), "t"], int_t = integrated_mean_DE[intersect(rownames(raw_mean_DE), rownames(integrated_mean_DE)), "t"])), aes(x = int_t, y = raw_t)) + 
  geom_point(size = 0.5, alpha = 0.3, color = "#7C81AD") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  ylim(c(0,10)) +
  xlab("|t score| of two-sample t-test (Integrated data)") + 
  ylab("|t score| of two-sample t-test (Log2 Raw UMI)") + 
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  xlim(c(0,10))

p_ttest_raw_vst = ggplot(data = abs(data.frame(raw_t = raw_mean_DE[intersect(rownames(raw_mean_DE), rownames(vst_mean_DE)), "t"], vst_t = vst_mean_DE[intersect(rownames(raw_mean_DE), rownames(vst_mean_DE)), "t"])), aes(x = vst_t, y = raw_t)) + 
  geom_point(size = 0.5, alpha = 0.3, color = "#7C81AD") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  ylim(c(0,10)) +
  xlab("|t score| of two-sample t-test (VST data)") + 
  ylab("|t score| of two-sample t-test (Log2 Raw UMI)") + 
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  xlim(c(0,10))

p_ttest_raw_cpm = ggplot(data = abs(data.frame(raw_t = raw_mean_DE[intersect(rownames(raw_mean_DE), rownames(cpm_mean_DE)), "t"], cpm_t = cpm_mean_DE[intersect(rownames(raw_mean_DE), rownames(cpm_mean_DE)), "t"])), aes(x = cpm_t, y = raw_t)) + 
  geom_point(size = 0.5, alpha = 0.3, color = "#7C81AD") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  ylim(c(0,10)) +
  xlab("|t score| of two-sample t-test (Log2 CPM data)") + 
  ylab("|t score| of two-sample t-test (Log2 Raw UMI)") + 
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  xlim(c(0,10))

ggarrange(p_ttest_raw_cpm, p_ttest_raw_vst, p_ttest_raw_integrated, ncol = 3)
```

# Number of DEGs from each method
```{r}
dflist = list(pois_glmm_df, 
              binom_glmm_df, 
              DESeq2.muscat,
              edgeR.cpm,
              MAST_donor_df, 
              Wilcox_Seurat_integrated_df,
              # MMdream,
              MMvst,
              MMpoisson)
titlelist = c("Poisson-glmm_UMI", 
              "Binomial-glmm_UMI",
              "pb-DESeq2_UMI", 
              "pb-edgeR_cpm",
              "MAST_CPM",
              "Wilcox_int",
              # "MMdream",
              "MMvst_VST",
              "MMpoisson_UMI")



number_input_hits = data.frame(DEGs = integer(0), Input = integer(0))
pois_glmm_inputgene = pois_glmm_df[[key]]$genes[!is.na(pois_glmm_df[[key]]$hits)]
for (df in dflist){
  number_input_hits = rbind(number_input_hits, 
                            c(sum(!is.na(df[[key]]$hits)), 
                              sum(df[[key]]$hits, na.rm = T))) 
}
colnames(number_input_hits) = c("Inputs", "DEGs")
number_input_hits$method = titlelist

number_input_hits = melt(number_input_hits, id = "method")
colnames(number_input_hits)[2:3] = c("Gene", "Number_of_genes")
```


```{r}
ggplot(data=number_input_hits, aes(x=factor(method, 
                                            levels = titlelist), y=Number_of_genes, fill = Gene)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label=Number_of_genes), vjust=-0.2, position=position_dodge(0.9), size = 2) +
  theme_bw() + theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  scale_fill_manual(values = c("#BCAAA4", "#5D4037")) +
  xlab("") +
  ylab("Number of genes") +
  ggtitle(paste("DEGs summary of ", key))
```


# Volcano plot
```{r,fig.width = 10, fig.height=6}
pvalcutoff = 0.05
v = 1.5
log2FCcutoff = log2(v)

p = list()
for (i in 1:length(dflist)){
  if(i %in% 6){
    xvar = 'avg_log2FC'
    yvar = 'p_val_adj'
  }else{
    xvar = 'log2FC'
    yvar = 'BH'
  }
  if (nrow(dflist[[i]][[key]])==0){next}
  p[[i]] = ggplot(na.omit(data.frame(dflist[[i]][[key]])), aes_string(x = xvar, y = sprintf("-log10(%s)", yvar), colour = "hits")) +
    geom_point(alpha = 0.5, size = 0.5) +
    labs(color = "hits") +
    geom_hline(yintercept=-log10(pvalcutoff),linetype=2) +
    geom_vline(xintercept=log2FCcutoff,linetype=2) + 
    geom_vline(xintercept=-log2FCcutoff,linetype=2) +
    ggtitle(titlelist[i]) + 
    theme_bw() + 
    theme(panel.grid = element_blank()) + 
    scale_color_manual(values = c("gray", "blue")) +
    labs(color = "DEGs") +
    xlim(c(-3,3)) + 
    ylim(c(0,50)) + 
    xlab("Log2 Fold Change") + 
    ylab("-Log10 (adj.pval)")
}

figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)

# Annotate the figure by adding a common labels
annotate_figure(figure,
                top = paste("Volcano plot",key),
                bottom = text_grob(paste( "log2FC cutoff ", paste0("log2(",v,")"),"; adj.pval cutoff", pvalcutoff)
                                   , x = 0.99, just = "right")
)
```


# Histogram of p-value/adj.p-value
```{r, fig.width = 10, fig.height=6}
p = list()
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  df = data.frame(genes = dflist[[i]][[key]]$genes)
  if(i %in% 6){
    df = cbind(df, dflist[[i]][[key]][,c("p_val", "p_val_adj")])
  }else{
    df = cbind(df, dflist[[i]][[key]][,c("pval", "BH")])
  }
  
  p[[i]] = ggplot(melt(df,id = "genes"), aes(x=value, color=variable)) +
geom_histogram(fill="white", alpha=0.5, position="identity") + theme_bw() + theme(panel.grid = element_blank()) + xlab(NULL) + ylab("Count") + coord_cartesian(ylim = c(0, 2000)) + ggtitle(titlelist[i]) + scale_colour_discrete(guide = guide_legend(title = NULL),labels = c("p-value", "adj.p-value"))
}
figure = ggarrange(plotlist = p, common.legend = TRUE, legend="bottom", nrow = 2, ncol = 4)

annotate_figure(figure, top = as.character(key))

```

# P-Value comparison across different methods
```{r,fig.width = 10, fig.height=6}
value_comparison("pval")
```

# Log2 fold change comparison across different methods
```{r,fig.width = 10, fig.height=6}
value_comparison("log2FC")
```

# Violin plot of log2mean of DEGs
```{r, fig.width=7}
p = list()
df = data.frame()
clusterlist = list(group1, group2)
clustername = c(paste(group1, collapse = "_"), paste(group2, collapse = "_"))
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  if (sum(dflist[[i]][[key]]$hits, na.rm = T)==0){next}
  hits_gene = dflist[[i]][[key]]$genes[which(dflist[[i]][[key]]$hits)]
  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:2){
    cellind = inputData$hippo_cluster %in% clusterlist[[j]]
    df = rbind(df, data.frame(Log2mean = log2(fun(assays(inputData)$counts[hits_gene,cellind])), Group = clustername[j], method = titlelist[i]))
  }
}
levels(df$Group) = clustername
```


```{r, fig.width=7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=Log2mean, fill = factor(Group, levels = levels(Group)))) +
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylab("Log2mean") + 
  ylim(c(-3,4)) + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) +
  labs(fill = "Group") +
  scale_fill_manual(values = hippo_palette[c(group1, group2)], breaks = levels(df$Group))

```

# Violin plot of gene expression frequency of DEGs
````{r, fig.width = 7}
p = list()
df = data.frame()
clusterlist = list(1:20, group1, group2)
clustername = c("All 20 \nclusters", c(paste(group1, collapse = "_"), paste(group2, collapse = "_")))
for (i in 1:length(dflist)){
  if (nrow(dflist[[i]][[key]])==0){next}
  if (sum(dflist[[i]][[key]]$hits, na.rm = T)==0){next}
  hits_gene = dflist[[i]][[key]]$genes[which(dflist[[i]][[key]]$hits)]
  if(length(hits_gene)>1){fun = rowMeans}else{fun = mean}
  for (j in 1:3){
    cellind = inputData$hippo_cluster %in% clusterlist[[j]]
    df = rbind(df, data.frame(freq = fun(assays(inputData)$counts[hits_gene,cellind]>0), cluster = clustername[j], method = titlelist[i]))
  }
}
levels(df$cluster) = clustername
```


````{r, fig.width = 7}
ggplot(df, aes(x = factor(method,levels = titlelist), y=freq, fill = factor(cluster, levels =  levels(cluster)))) + 
  geom_violin() + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  ylab("Gene expression frequency") + 
  ylim(c(0,1)) + 
  xlab("") + 
  scale_fill_manual(values = c("gray", hippo_palette[c(group1, group2)]), breaks = levels(df$cluster)) +
  labs(fill = "Group") +
  theme(axis.text.x = element_text(angle = 20, margin = margin(t = 15))) 
```


# Heatmap of top DEGs

## Poisson-glmm DEGs

### UMI counts
```{r}
sort_log2FC = sort(abs(pois_glmm_df[[key]]$log2FC[pois_glmm_df[[key]]$hits]), index.return = T, decreasing = T)
hitgenes = pois_glmm_df[[key]]$genes[which(pois_glmm_df[[key]]$hits)][sort_log2FC$ix]
hits_in_int = hitgenes[hitgenes%in%rownames(subgroupsce_integrated)]
hits_notin_int = na.omit(hitgenes[!hitgenes%in%rownames(subgroupsce_integrated)])
mat = assays(subgroupsce)$counts[c(hits_in_int, hits_notin_int),]
#row annotation
annotation_log2FC = data.frame(log2FC = abs(pois_glmm_df[[key]]$log2FC[match(c(hits_in_int, hits_notin_int) ,pois_glmm_df[[key]]$genes)]), `Integrated data` = c(rep("Incluede", length(hits_in_int)), rep("Not included", length(hits_notin_int))))
rownames(annotation_log2FC) = rownames(mat)
#col annotation
annotation_df = data.frame(Donor = subgroupsce$donor, Group = factor(subgroupsce$hippo_cluster, levels = c(group1, group2)))
rownames(annotation_df) = colnames(mat)
annotation_df = annotation_df[with(annotation_df, order(Group, Donor)), ]
# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Poisson-glmm DEGs\nin UMI counts (", nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```

### VST data
```{r}
mat = vstcounts[c(hits_in_int, hits_notin_int), subset_ind]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Poisson-glmm DEGs\nin VST data (", nrow(mat)," DEGs)"),
         annotation_col = annotation_df,
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```


### CPM data
```{r}
mat = assays(subgroupsce)$cpm[c(hits_in_int, hits_notin_int),]/1000

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Poisson-glmm DEGs\nin CPM data (", nrow(mat)," DEGs)"),
         annotation_col = annotation_df,
         annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```

### Integrated data
```{r}
mat = assays(subgroupsce_integrated)$counts[hits_in_int,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Poisson-glmm DEGs\nin integrated data (", length(hits_in_int)," DEGs)"),
         annotation_col = annotation_df,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))


```

## Additional DEGs from other methods

### pb-DESeq2
```{r}
sort_log2FC = sort(abs(DESeq2.muscat[[key]]$log2FC[DESeq2.muscat[[key]]$hits]), index.return = T, decreasing = T)
hitgenes_deseq2 = DESeq2.muscat[[key]]$genes[which(DESeq2.muscat[[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_deseq2[!hitgenes_deseq2%in%hitgenes])
mat = assays(subgroupsce)$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("pb-DESeq2 DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         # annotation_row = annotation_log2FC,
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```

### Binomial-glmm
```{r}
sort_log2FC = sort(abs(binom_glmm_df[[key]]$log2FC[binom_glmm_df[[key]]$hits]), index.return = T, decreasing = T)
hitgenes_binom = binom_glmm_df[[key]]$genes[which(binom_glmm_df[[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_binom[!hitgenes_binom%in%hitgenes])
mat = assays(subgroupsce)$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Binomial-glmm DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```

### MAST
```{r}
sort_log2FC = sort(abs(MAST_donor_df[[key]]$log2FC[MAST_donor_df[[key]]$hits]), index.return = T, decreasing = T)
hitgenes_MAST = MAST_donor_df[[key]]$genes[which(MAST_donor_df[[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_MAST[!hitgenes_MAST%in%hitgenes])
mat = assays(subgroupsce)$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("MAST DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```


### MMpoisson
```{r}
sort_log2FC = sort(abs(MMpoisson[[key]]$log2FC[MMpoisson[[key]]$hits]), index.return = T, decreasing = T)
hitgenes_MMpoisson = MMpoisson[[key]]$genes[which(MMpoisson[[key]]$hits)][sort_log2FC$ix]
hits_notin_glmm = na.omit(hitgenes_MMpoisson[!hitgenes_MMpoisson%in%hitgenes])
mat = assays(subgroupsce)$counts[hits_notin_glmm,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("MMpoisson DEGs \nnot identified by Poisson-glmm (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))
```

## DEGs in Poisson-glmm not identified by MMpoisson

In the MMpoisson model, cell type is considered as a random effect. This approach treats certain aspects of cell type variations as random factors. Consequently, it may obscure the true variation in cell types, limiting its ability to accurately reveal the specific differences between different cell types.   

Additionally, the library size is employed as an offset to normalize the counts. That is, the model is considering rate instead of counts. Suppose some genes are highly expressed in one cell type than the other, the absolute difference could be eliminate after accounting for library size. This normalization approach may inadvertently mask certain gene expression differences between cell types.

```{r}
hitgenes_MMpoisson = MMpoisson[[key]]$genes[which(MMpoisson[[key]]$hits)]
hits_notin_MMpoisson = na.omit(hitgenes[!hitgenes%in%hitgenes_MMpoisson])
mat = assays(subgroupsce)$counts[hits_notin_MMpoisson,]

# cell level
pheatmap(mat[,rownames(annotation_df)], 
         main = paste0("Poisson-glmm DEGs \nnot identified by MMpoisson (",nrow(mat)," DEGs)"),
         annotation_col = annotation_df, 
         annotation_colors = annotation_colors,
         cluster_rows=F, cluster_cols=F, 
         show_colnames = F, show_rownames = F,
         annotation_names_row = F,
         color=colorRampPalette(c("navy", "white", "red"))(10), breaks = seq(0,3, length.out = 10))

```



# MA plot 
```{r, fig.width = 10, fig.height=5}
fig_FC_mean = FC_mean_plot(subgroupsce@metadata$log2mean, c(-6,6), "Log2 mean") 
annotate_figure(fig_FC_mean, top = paste("Log2FC vs. Log2 mean ",key))
```

# Enrichment analysis

## GO object
```{r, fig.width = 6, fig.height=8}
original_gene_list = pois_glmm_df[[key]]$log2FC
names(original_gene_list) = pois_glmm_df[[key]]$genes
gene_list = na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
genes = pois_glmm_df[[key]]$genes[which(pois_glmm_df[[key]]$hits)]
# create GO object
go_enrich = enrichGO(gene = genes,
                      universe = names(gene_list),
                      OrgDb = "org.Hs.eg.db", 
                      keyType = 'SYMBOL',
                      readable = T,
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)
```
```{r, fig.width = 8, fig.height=6}
dotplot(go_enrich)
```
```{r, fig.width = 14, fig.height=7}
cnetplot(go_enrich, categorySize="pvalue", foldChange=gene_list)
```


## enrichKEGG object
```{r, fig.width = 10, fig.height=5,}
# Convert gene IDs for enrichKEGG function
# We will lose some genes here because not all IDs will be converted
ids = bitr(names(original_gene_list), fromType = "SYMBOL", toType = "ENTREZID", OrgDb="org.Hs.eg.db") 
dedup_ids = ids[!duplicated(ids[c("SYMBOL")]),]

# Create a vector of the gene universe
match_logic = pois_glmm_df[[key]]$genes %in% dedup_ids$SYMBOL
kegg_gene_list <- pois_glmm_df[[key]]$log2FC[match_logic]
names(kegg_gene_list) <- dedup_ids$ENTREZID
kegg_gene_list<-na.omit(kegg_gene_list)
kegg_gene_list = sort(kegg_gene_list, decreasing = TRUE)
# hits
kegg_genes = na.omit(dedup_ids$ENTREZID[pois_glmm_df[[key]]$hits[match_logic]])
# create enrichKEGG object
kk = enrichKEGG(gene=kegg_genes,
                universe=names(kegg_gene_list),
                organism="hsa", pvalueCutoff = 0.05, keyType = "ncbi-geneid")
```
```{r,fig.width = 8, fig.height=4}
dotplot(kk, 
        showCategory = 10, 
        title = "Enriched Pathways",
        font.size = 8)

```

```{r,fig.width = 10, fig.height=5}
cnetplot(kk, categorySize="pvalue", foldChange=kegg_gene_list)
```